The weight of the mountains: Constraints on tectonic stress, friction, and fluid pressure in the 2008 Wenchuan earthquake from estimates of topographic loading

### 2014-09-17 revisions/rebuttal.md

>In general, the models share a common geometry and pattern of slip distribution
above 10 km or so; i.e. the locations of high and low slip patches and the slip
rake are similar in all of the models. Though some of models use discrete fault
segments for the Beichuan fault, whereas others use a single continuous fault,
the fault planes are essentially collocated. However, there is significant
variability in the magnitude of slip in the different models; for example, the
maximum slip magnitude in the Qi 'rough' model is about twice that of the
Feng model. Large differences also exist in the deeper geometries of the
models: the Qi, Fielding and Shen models all have horizontal or subhorizontal
thrust flats at or below 15 km, though only minor slip is assumed to have
occurred on these fault segments. In contrast, the Feng and Zhang models are
planar to their lower end at 25-30 km.

The significance of the findings highlighted here by the reviewer are discussed
in the discussion (Section 6.4, particularly lines 504-511). In general, our
chosen expository style in this paper is more geared towards description of the
results rather than (over-)interpretation. The results themselves are described,
and the results files are made available for readers who are interested in
probing them for themselves. Because our results cover both topographic and
tectonic stresses, as well as fault strength parameters, there are a lot of
implications for anything from earthquake processes to orogenic development,
and we feel it is better to be reasonably brief in our discussion of 
individual aspects of these implications.

@@

## R11
> The introduction of the study says the work intends to "investigate the
> effects of topographic stress on the faults" (line 48) and then "bracket the
> tectonic stress field, pore fluid pressure, and static friction of the fault"
> (line54). However, in beginning to describe the topographic stress resolved on
> faults, section 3 refers to an upcoming manuscript for analysis of how these
> stresses might affect faulting. This is indeed an interesting aspect of the
> study, but either that information should be included here or the focus on
> influencing rupture process should be removed from the abstract and
> introduction.

There is nothing in the abstract that references a topographic influence on
rupture processes. In one sentence in the introduction (lines 41-45), we write
that topographic stresses *may* influence rupture propagation. We do not
'focus' on this. Additionally, as promised in the two quoted passages (from
lines 48 and 54), we do investigate the effects of topographic stress on the
faults, and bracket the tectonic stresses, friction, and fluid pressure on the
faults. This is the focus of the paper.

The first-order 'effects of topographic stress on the faults' is whether those
stresses promote or inhibit slip on the faults. We show that topographic
stresses are overwhelmingly resistive/inhibiting to fault slip, both by adding
shear stress to the faults in the opposite direction to slip, and by adding
significant normal stress to the faults.

The influence of topographic stresses on a rupture that does occur may be a
second-order effect. We see that topographic stresses, particularly normal
stress, is negatively correlated with slip magnitude, and we discuss this in
lines 248-255. However, any real discussion on this other than the observation
that it occurs is beyond what we can comfortably say at this time, without
being more speculative than we would wish. However, we believe it is an
important observation, and it is certainly one which we have already made, and
so it would be inappropriate to sweep it under the rug because we do not yet
understand it at publishable level. It may be that other researchers are
interested, or have better means to investigate it (i.e. they are experts in
dynamic rupture modeling), and we should not inhibit them.

## R12
> Overall, the manuscript, though well organized, seems to have an unbalanced
> emphasis. The specific details of the analytical methods are present, but
> overall the results should be more fully described and the discussion more
> fully developed.

This is definitely a 'methods' paper, and this work was performed with the
broader goal of working to quantify tectonic stresses in a larger sense. We
chose this earthquake to start with because it's well-documented and has a
complex coseismic slip distribution, and because it lies at the base of a very
large plateau that has been suggested to be expanding outward due to
topographic stresses. We looked into these effects, framed the investigation in
terms of questions (e.g., do topographic stresses promote or inhibit the
observed slip on the faults), and answered them.

Nonetheless, the results and the discussion make up the bulk of the paper. The
methods sections (2 and 4) are 6.5 'manuscript' pages in sum, which includes
many equations that take up a lot of space. The results sections (3 and 5) are
about 2 pages, and 6 of the 10 figures. The discussion is 8 pages, by far the
longest section.

@@

## R13
> Line 24: Though I can see the appeal of the selected quotation, it would be
more appropriate for an oral presentation or a monograph on the subject. In
this context, it comes across as awkward and out of place.

We have removed it.

## R14
> Section 1.1: additional relevant references to consider: Fialko, Rivera, and
Kanamori (2005 GJI), Ghosh et al (2006 Geology), Flesch et al (2007 GJI)

We have added a reference to the Fialko paper. The Flesch and Ghosh papers
mentioned here are 'thin viscous sheet' type papers, similar to those already
mentioned in the review (including a later paper by Flesch); those types of
studies are of a much larger (orogen to global) scale than our study, and
both a longer list of studies and a more detailed discussion of the studies
is a bit out of the scope of the current paper.

@@

## R15
> Lines 224-230 describe the magnitudes of stress tensor components, but this
would be better explained with a figure. If not as its own figure, could Mzz at
least be incorporated into figure 3?

We have added vertical topographic stress to Figures 3 and 9.

@@

## R16
> Section 6.0 contrasts the methods of this study to those of Luttrell et al.
2011 study of the Maule earthquake. This is an understandable comparison, but
because the Wenchuan earthquake includes both strike-slip and dip-slip
components, it is actually more similar to the Luttrell and Sandwell (2012 JGR)
study of Mid Ocean Ridges that includes both strike-slip and normal earthquakes
as constraints.

Actually, we feel that our manuscript is more similar to the Luttrell et al.
2011 paper because that paper uses a coseismic slip model in the stress
inversion, and because it involves a single event. The Luttrell and Sandwell
paper uses focal mechanisms, which are basically point sources, and considers
multiple earthquakes. This is a bit more different with respect to our paper.
## R17
> The introduction lists 7 coseismic slip models of the Wenchuan earthquake (line 124), and then the analysis refers to 5 of them (line 213). As the authors noticed, all these models essentially draw from the same set of geodetic data, and all have essentially the same result, and thus all give this study essentially the same result (line 343). Is there a reason these particular models were chosen for analysis? Why not just choose one, or if they truly are different enough to influence these results, why not use them all? Some justification of this choice is warranted.

We used all of the models that we could find that were available (i.e. the slip
model files themselves were published), had proper georeferencing, and seemed
to be of sufficiently high quality. There are many models that did not attempt
a good geometric fit to the available data--e.g. they consist of a single,
planar fault segment--or models with very poor coseismic slip resolution. We
did not feel it wise to include these (garbage in equals garbage out, as they
say). Additionally, we did not feel it wise to list models that we thought
were of too low quality.

We have added some discussion of the models used, and the criteria for
including them, to Section 2.3, as described in R10.

## R18
> --Typos

> Lines 177 and 182: is z positive up or down? Clarify this

> Line 258: earthquake, not fault, initiated

> Line 274: "are" -> "and"? complete sentence with period

> Line 509: typo, delete "to close"

> Line 556: If you acknowledge some of your data sources, you should acknowledge all of your data sources (e.g., scientific drilling, WSM, etc)

> Fig 3 and 9 need 33ºN label correction

Corrected.
